Carbonate chemistry at Ny-Ålesund

### 2022-03-30 awipev-CO2.Rmd
refined pangaea and saving nuts fig

```{r impact of nutrients, echo=FALSE, warning=FALSE, message=FALSE}
# boxplot
ts_po4_bp <- d_all %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = po4)) +
  #geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE, na.rm = TRUE) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3, na.rm = TRUE) +
  labs(y = "po4", x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())

ts_Si_bp <- d_all %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = si)) +
  #geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE, na.rm = TRUE) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3, na.rm = TRUE) +
  labs(y = "si", x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())
ts_po4_bp + ts_Si_bp
# Effects of changes in po4 and si
# 1st cakculate mean values
means <- d_all %>% 
  group_by(montht) %>% 
  summarize(mean_ct = mean(ct, na.rm = TRUE),
            mean_at_calc = mean(at_calc, na.rm = TRUE),
            mean_pco2_calc = mean(pco2_calc_fb, na.rm = TRUE),
            mean_sal = mean(sal_fb_mix, na.rm = TRUE),
            mean_temp = mean(temp_fb, na.rm = TRUE)) %>% 
  ungroup()
mean_ct <-  mean(means$mean_ct)
mean_at_calc <-  mean(means$mean_at_calc)
mean_pco2_calc <-  mean(means$mean_pco2_calc)
mean_sal <-  mean(means$mean_sal)
mean_temp <-  mean(means$mean_temp)
# then check impacts of 0, min and max nutrient values using pco2 and AT
carb_po4_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = c(0, min(d_all$po4/1e6, na.rm = TRUE), max(d_all$po4/1e6, na.rm = TRUE)), Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_po4_0a <- carb_po4_1[1,] - carb_po4_1[3,] # diff between no PO4 and max PO4
diff_po4_1a <- carb_po4_1[2,] - carb_po4_1[3,] # diff between min PO4 and max PO4
carb_Si_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = c(0, min(d_all$si/1e6, na.rm = TRUE), max(d_all$si/1e6, na.rm = TRUE)), 
                k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_Si_0a <- carb_Si_1[1,] - carb_Si_1[3,] # diff between no Si and max Si
diff_Si_1a <- carb_Si_1[2,] - carb_Si_1[3,] # diff between no Si and max Si
# then check impacts of O, min and max nutrient values using AT-DIC pair
carb_po4_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = c(0, min(d_all$po4/1e6, na.rm = TRUE), max(d_all$po4/1e6, na.rm = TRUE)), Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_po4_0b <- carb_po4_2[1,] - carb_po4_2[3,] # diff between no PO4 and max PO4
diff_po4_1b <- carb_po4_2[2,] - carb_po4_2[3,] # diff between min PO4 and max PO4
carb_Si_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = c(0, min(d_all$si/1e6, na.rm = TRUE), max(d_all$si/1e6, na.rm = TRUE)), 
                k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_Si_0b <- carb_Si_2[1,] - carb_Si_2[3,] # diff between no Si and max Si
diff_Si_1b <- carb_Si_2[2,] - carb_Si_2[3,] # diff between no Si and max Si
# then check impacts of O, max PO4 and NO3 values using pCO2 and AT as well as AT-DIC pair
carb_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = c(0, max(d_all$po4/1e6, na.rm = TRUE)), Sit = c(0, max(d_all$si/1e6, na.rm = TRUE)),
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = c(0, max(d_all$po4/1e6, na.rm = TRUE)), Sit = c(0, max(d_all$si/1e6, na.rm = TRUE)),
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_max_0a <- carb_1[1,] - carb_1[2,] # pco2-AT: diff between (0 PO4 and Si) and (max PO4 and Si)
diff_max_0b <- carb_2[1,] - carb_2[2,] # AT-CT: diff between (0 PO4 and Si) and (max PO4 and Si)

@@

pangaea <- as_tibble(d_all) %>% 
  dplyr::select(datetime, depth, sal_insitu_ctd, sal_fb, temp_insitu_11m, temp_insitu_ctd, temp_fb, temp_sf, at, ct, ph_s_sf_temp_insi, ph_s_dur_temp_fb, ph_calc_fb, pco2_fb_insi_temp, pco2_calc_fb_insi_temp, at_calc, pH_dur, pHint_tot_sf) %>%
  dplyr::rename(`pressure [dbar]` = depth,
                `s_insitu [unit]` = sal_insitu_ctd,
                `s_fb [unit]` = sal_fb, 
                `t_11m [°C]` = temp_insitu_11m,
                `t_ctd [°C]` = temp_insitu_ctd, 
                `t_fb [°C]` = temp_fb, 
                `t_s [°C]` = temp_fb,
                `t_sf [°C]` = temp_sf, 
                `at [umol kg-1]` = at, 
                `at_calc [umol kg-1]` = at_calc,
                `ct [umol kg-1]` = ct, 
                `ph_s_sf_t_insi [total scale]` = ph_s_sf_temp_insi, 
                `ph_s_dur_t_fb [total scale]` = ph_s_dur_temp_fb,
                `pco2 [uatm]` = pco2_fb_insi_temp,
                `pco2_calc [uatm]` = pco2_calc_fb_insi_temp,
                `pH_sf [total scale]` = pHint_tot_sf,
                `pH_dur [total scale]` = pH_dur,
                `pH_calc [total scale]` = ph_calc_fb
                )
saveRDS(file= paste0(path, "fb_data/pangaea.rds"), pangaea)
write_csv(pangaea, file = paste0(path, "fb_data/pangaea.csv"))
# Metadata Pangaea
# pressure = pressure in situ (profiler)
# s_insitu = salinity in situ (profiler)
# s_fb = salinity in FerryBox 
# t_11m = temperature in situ at 11 m
# t_ctd = temperature in situ (profiler), 
# t_fb = temperature in FerryBox 
# t_sf = temperature SeaFET (profiler)
# at = total alkalinity, discrete measurements (FerryBox)
# at_calc = total alkalinity, calculated from salinity
# ct = dissolved inorganic carbon, discrete measurements (FerryBox)
# ph_s_sf_t_insi = spectrophotometric pH in situ at in situ temperature 
# ph_s_dur_t_fb = spectrophotometric pH in FerryBox at FerryBox temperature
# pco2 = pCO2 in Ferrybox expressed at in situ temperature
# pco2_calc = pCO2 in the FerryBox calculated from discrete measurements of AT and CT and expressed at in situ temperature
# pH_sf = pH seaFET at in situ temperature (profiler)
# ph_dur = pH Durafet at FerryBox temperature (FerryBox)
# ph_calc = pH at FerryBox temperature calculated from CT and AT in the FerryBox
```

>>>>

```{r impact of nutrients, echo=FALSE, warning=FALSE, message=FALSE}
# boxplot
ts_po4_bp <- d_all %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = po4)) +
  #geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE, na.rm = TRUE) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3, na.rm = TRUE) +
  labs(y = "PO4", x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())

ts_Si_bp <- d_all %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = si)) +
  #geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE, na.rm = TRUE) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3, na.rm = TRUE) +
  labs(y = "Si", x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())
p <- ts_po4_bp + ts_Si_bp
ggsave(filename = "figures/nutrients.png", plot = p, width = 25, height = 20, units = "cm")

# Effects of changes in po4 and si
# 1st cakculate mean values
means <- d_all %>% 
  group_by(montht) %>% 
  summarize(mean_ct = mean(ct, na.rm = TRUE),
            mean_at_calc = mean(at_calc, na.rm = TRUE),
            mean_pco2_calc = mean(pco2_calc_fb, na.rm = TRUE),
            mean_sal = mean(sal_fb_mix, na.rm = TRUE),
            mean_temp = mean(temp_fb, na.rm = TRUE)) %>% 
  ungroup()
mean_ct <-  mean(means$mean_ct)
mean_at_calc <-  mean(means$mean_at_calc)
mean_pco2_calc <-  mean(means$mean_pco2_calc)
mean_sal <-  mean(means$mean_sal)
mean_temp <-  mean(means$mean_temp)
# then check impacts of 0, min and max nutrient values using pco2 and AT
carb_po4_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = c(0, min(d_all$po4/1e6, na.rm = TRUE), max(d_all$po4/1e6, na.rm = TRUE)), Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_po4_0a <- carb_po4_1[1,] - carb_po4_1[3,] # diff between no PO4 and max PO4
diff_po4_1a <- carb_po4_1[2,] - carb_po4_1[3,] # diff between min PO4 and max PO4
carb_Si_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = c(0, min(d_all$si/1e6, na.rm = TRUE), max(d_all$si/1e6, na.rm = TRUE)), 
                k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_Si_0a <- carb_Si_1[1,] - carb_Si_1[3,] # diff between no Si and max Si
diff_Si_1a <- carb_Si_1[2,] - carb_Si_1[3,] # diff between no Si and max Si
# then check impacts of O, min and max nutrient values using AT-DIC pair
carb_po4_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = c(0, min(d_all$po4/1e6, na.rm = TRUE), max(d_all$po4/1e6, na.rm = TRUE)), Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_po4_0b <- carb_po4_2[1,] - carb_po4_2[3,] # diff between no PO4 and max PO4
diff_po4_1b <- carb_po4_2[2,] - carb_po4_2[3,] # diff between min PO4 and max PO4
carb_Si_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = c(0, min(d_all$si/1e6, na.rm = TRUE), max(d_all$si/1e6, na.rm = TRUE)), 
                k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_Si_0b <- carb_Si_2[1,] - carb_Si_2[3,] # diff between no Si and max Si
diff_Si_1b <- carb_Si_2[2,] - carb_Si_2[3,] # diff between no Si and max Si
# then check impacts of O, max PO4 and NO3 values using pCO2 and AT as well as AT-DIC pair
carb_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = c(0, max(d_all$po4/1e6, na.rm = TRUE)), Sit = c(0, max(d_all$si/1e6, na.rm = TRUE)),
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = c(0, max(d_all$po4/1e6, na.rm = TRUE)), Sit = c(0, max(d_all$si/1e6, na.rm = TRUE)),
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_max_0a <- carb_1[1,] - carb_1[2,] # pco2-AT: diff between (0 PO4 and Si) and (max PO4 and Si)
diff_max_0b <- carb_2[1,] - carb_2[2,] # AT-CT: diff between (0 PO4 and Si) and (max PO4 and Si)

@@

pangaea <- as_tibble(d_all) %>% 
  dplyr::select(datetime, depth, sal_insitu_ctd, sal_fb, temp_insitu_11m, temp_insitu_ctd, temp_fb, temp_sf, at, ct, ph_s_sf_temp_insi, ph_s_dur_temp_fb, ph_calc_fb, pco2_fb_insi_temp, pco2_calc_fb_insi_temp, at_calc, pH_dur, pHint_tot_sf) %>%
  dplyr::rename(`pressure [dbar]` = depth,
                `s_insitu [unit]` = sal_insitu_ctd,
                `s_fb [unit]` = sal_fb, 
                `t_11m [°C]` = temp_insitu_11m,
                `t_ctd [°C]` = temp_insitu_ctd, 
                `t_fb [°C]` = temp_fb, 
                `t_sf [°C]` = temp_sf, 
                `at [umol kg-1]` = at, 
                `at_calc [umol kg-1]` = at_calc,
                `ct [umol kg-1]` = ct, 
                `ph_s_sf_t_insi [total scale]` = ph_s_sf_temp_insi, 
                `ph_s_dur_t_fb [total scale]` = ph_s_dur_temp_fb,
                `pco2 [uatm]` = pco2_fb_insi_temp,
                `pco2_calc [uatm]` = pco2_calc_fb_insi_temp,
                `pH_sf [total scale]` = pHint_tot_sf,
                `pH_dur [total scale]` = pH_dur,
                `pH_calc [total scale]` = ph_calc_fb
                )
saveRDS(file= paste0(path, "fb_data/pangaea.rds"), pangaea)
write_csv(pangaea, file = paste0(path, "fb_data/pangaea.csv"))
# Metadata Pangaea
# pressure = pressure in situ (profiler)
# s_insitu = salinity in situ (profiler)
# s_fb = salinity (static at 11 m) 
# t_11m = temperature in situ (static at 11 m)
# t_ctd = temperature in situ (profiler), 
# t_fb = temperature (FerryBox)
# t_sf = temperature SeaFET (profiler)
# at = total alkalinity, discrete measurements (FerryBox)
# at_calc = total alkalinity, calculated from salinity
# ct = dissolved inorganic carbon, discrete measurements (FerryBox)
# ph_s_sf_t_insi = spectrophotometric pH in situ at in situ temperature (profiler)
# ph_s_dur_t_fb = spectrophotometric pH in FerryBox at FerryBox temperature (FerryBox)
# pco2 = pCO2 in Ferrybox expressed at in situ temperature (FerryBox)
# pco2_calc = pCO2 in the FerryBox calculated from discrete measurements of AT and CT and expressed at in situ temperature
# pH_sf = pH seaFET at in situ temperature (profiler)
# ph_dur = pH Durafet at FerryBox temperature (FerryBox)
# ph_calc = pH at FerryBox temperature calculated from CT and AT in the FerryBox
```
